---
title: "Root shoot ratio in DGVMs"
author: "Benjamin D. Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
---

```{r include=FALSE}
library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(readr)
library(stringr)
library(rbeni)

filnams <- read_csv( "filnams_trendy_v8_S2.csv" )
varnams <- read_csv( "varnams_trendy_v8_S2.csv" )

# modeltype <- read_csv( "modeltype_trendy_v8.csv" ) %>% 
  # mutate( col=ifelse( cn==1, "royalblue3", "springgreen3" ) )
```

## Process TRENDY v8 files

From downloaded TRENDY outputs (here, simulation S2: only CO2 is changing), get global fields of mass fractions $M_R, M_F, M_W$ and root:shoot ratio $(M_F + M_W)/M_R$ as the mean over several years. This is implemented in a Bash script.

### Get forest mask

Use >10% tree cover cutoff to determine forest mask. Information about tree vs. grass-PFT coverage needs to be extracted from NetCDF files for each model separately. This is done by hand and information saved in `pft_tree_grass_trendy_v8.csv`.

Ingest information about tree-PFT indeces.
```{r}
df_treegrass <- read_csv("pft_tree_grass_trendy_v8.csv")

## Get tree-indices as a list for each model (nicely reading from CSV file)
get_tree_pft_id <- function(df){
  df %>% 
    dplyr::select(3:ncol(.)) %>% 
    pivot_longer(1:ncol(.), names_to = "pft_id", values_to = "pft_type") %>% 
    mutate(pft_id = as.integer(pft_id)) %>% 
    filter(pft_type == "Tree") %>% 
    pull(pft_id)
}

list_tree_pft_id <- purrr::map(as.list(seq(nrow(df_treegrass))), ~get_tree_pft_id(slice(df_treegrass, .)))
names(list_tree_pft_id) <- df_treegrass %>% pull(modl)
```

Get temporal mean PFT cover (using CDO timemean)
```{r}
df_nc <- df_treegrass %>% 
  dplyr::filter(!is.na(Filename)) %>% 
  dplyr::select(modl, Filename) %>% 
  mutate(filnam_base = stringr::str_remove(Filename, ".nc")) %>% 
  mutate(dir = "~/data/trendy/v8/") %>% 
  mutate(cmd = paste("./get_timmean.sh", filnam_base, dir))
```

Execute function using CDO.
```{r eval=FALSE}
out <- purrr::map(as.list(df_nc %>% pull(cmd)), ~system(., intern = TRUE))
```

read PFT cover data for each model (annual mean files)
```{r}
dir <- "~/data/trendy/v8/"
df_nc <- df_nc %>% 
  mutate(filnam_mean = paste0(dir, filnam_base, "_MEAN.nc")) %>% 
  mutate(nc = purrr::map(filnam_mean, ~read_nc_onefile(.)))
```

Sum over tree PFTs for each model individually
```{r}
modl <- "CABLE-POP"
idx <- which(df_nc$modl == modl)
nc <- df_nc$nc[[idx]]
nc$vars$landCoverFrac %>% dim()
nc$vars$landCoverFrac <- apply( nc$vars$landCoverFrac[,, list_tree_pft_id[[modl]] ], MARGIN = c(1,2), FUN = sum )
df_nc$nc_tree[[idx]] <- nc

modl <- "CLASS-CTEM"
idx <- which(df_nc$modl == modl)
nc <- df_nc$nc[[idx]]
nc$vars$landCoverFrac %>% dim()
nc$vars$landCoverFrac <- apply( nc$vars$landCoverFrac[,, list_tree_pft_id[[modl]] ], MARGIN = c(1,2), FUN = sum )
df_nc$nc_tree[[idx]] <- nc

modl <- "ISAM"
idx <- which(df_nc$modl == modl)
nc <- df_nc$nc[[idx]]
nc$vars$landCoverFrac %>% dim()
nc$vars$landCoverFrac <- apply( nc$vars$landCoverFrac[,, list_tree_pft_id[[modl]] ], MARGIN = c(1,2), FUN = sum )
df_nc$nc_tree[[idx]] <- nc

modl <- "JSBACH"
idx <- which(df_nc$modl == modl)
nc <- df_nc$nc[[idx]]
nc$vars$landCoverFrac %>% dim()
nc$vars$landCoverFrac <- apply( nc$vars$landCoverFrac[list_tree_pft_id[[modl]],, ], MARGIN = c(2,3), FUN = sum )
df_nc$nc_tree[[idx]] <- nc

## info not available for JULES
modl <- "JULES-ES-1.0"
idx <- which(df_nc$modl == modl)
nc <- df_nc$nc[[idx]]
nc$vars$landCoverFrac %>% dim()
# nc$vars$landCoverFrac <- apply( nc$vars$landCoverFrac[,, list_tree_pft_id[[modl]] ], MARGIN = c(1,2), FUN = sum )
df_nc$nc_tree[[idx]] <- NA

modl <- "LPJ-GUESS"
idx <- which(df_nc$modl == modl)
nc <- df_nc$nc[[idx]]
nc$vars$landCoverFrac %>% dim()
nc$vars$landCoverFrac <- apply( nc$vars$landCoverFrac[ list_tree_pft_id[[modl]],,], MARGIN = c(2,3), FUN = sum )
df_nc$nc_tree[[idx]] <- nc

modl <- "LPX-Bern"
idx <- which(df_nc$modl == modl)
nc <- df_nc$nc[[idx]]
nc$vars$landCoverFrac %>% dim()
nc$vars$landCoverFrac <- apply( nc$vars$landCoverFrac[list_tree_pft_id[[modl]],,], MARGIN = c(2,3), FUN = sum )
df_nc$nc_tree[[idx]] <- nc

modl <- "ORCHIDEE"
idx <- which(df_nc$modl == modl)
nc <- df_nc$nc[[idx]]
nc$vars$landCoverFrac %>% dim()
nc$vars$landCoverFrac <- apply( nc$vars$landCoverFrac[,, list_tree_pft_id[[modl]] ], MARGIN = c(1,2), FUN = sum )
nc$lat <- rev(nc$lat) ## flip latitude
df_nc$nc_tree[[idx]] <- nc

modl <- "ORCHIDEE-CNP"
idx <- which(df_nc$modl == modl)
nc <- df_nc$nc[[idx]]
nc$vars$landCoverFrac %>% dim()
nc$vars$landCoverFrac <- apply( nc$vars$landCoverFrac[,, list_tree_pft_id[[modl]] ], MARGIN = c(1,2), FUN = sum )
nc$lat <- rev(nc$lat) ## flip latitude
df_nc$nc_tree[[idx]] <- nc

modl <- "SDGVM"
idx <- which(df_nc$modl == modl)
nc <- df_nc$nc[[idx]]
nc$vars$landCoverFrac %>% dim()
nc$vars$landCoverFrac <- apply( nc$vars$landCoverFrac[,, list_tree_pft_id[[modl]] ], MARGIN = c(1,2), FUN = sum )
df_nc$nc_tree[[idx]] <- nc

modl <- "VISIT"
idx <- which(df_nc$modl == modl)
nc <- df_nc$nc[[idx]]
nc$vars$landCoverFrac %>% dim()
nc$vars$landCoverFrac <- apply( nc$vars$landCoverFrac[,, list_tree_pft_id[[modl]] ], MARGIN = c(1,2), FUN = sum )
df_nc$nc_tree[[idx]] <- nc
```

Look at some data.
```{r}
df_nc$nc_tree[[11]] %>% plot_map3(varnam = "landCoverFrac")
```

Sum over tree PFTs
```{r}
sum_tree_pft_cover <- function(nc, tree_ids){
  nc$vars$landCoverFrac <- apply(nc$vars$landCoverFrac[,,tree_ids,], MARGIN = c(1,2,4), FUN = sum)
  return(nc)
}
list_nc <- list()
for (imodl in (df_nc %>% pull(modl))){
  df_tmp <- df_nc %>% filter(modl == imodl)
  list_nc[[imodl]] <- sum_tree_pft_cover(
    df_tmp$nc,
    list_tree_pft_id[[imodl]]
  )
}
names(list_nc) <- df_nc %>% pull(modl)
```

### Wood mass fraction

Determine for which models we have all the files (cWood and cVeg) to calculate wood mass fractions and create system command.
```{r}
df <- filnams %>% 
  filter(!is.na(cWood)==1 & !is.na(cVeg)==1) %>% 
  select(modl, filn_cWood = cWood, filn_cVeg = cVeg) %>% 
  left_join(
    varnams %>% 
      select(modl, varn_cWood = cWood, varn_cVeg = cVeg),
    by = "modl"
  ) %>% 
  mutate_at(vars(starts_with("filn_")), ~str_replace(., ".nc", "")) %>% 
  mutate(dir = paste0("~/data/trendy/v8/")) %>% 
  mutate(cmd = paste("./calc_mass_fraction.sh", filn_cWood, filn_cVeg, dir))
```

Execute function using CDO.
```{r}
out <- purrr::map(as.list(df %>% pull(cmd)), ~system(., intern = TRUE))
```

### Leaf mass fraction

Determine for which models we have all the files (cWood and cVeg) to calculate wood mass fractions and create system command.
```{r}
modls_ml <- availvars %>% 
  filter(cLeaf==1 & cVeg==1) %>% 
  pull(modl)

df <- filnams %>% 
  filter(modl %in% modls_ml) %>% 
  select(modl, filn_cLeaf = cLeaf, filn_cVeg = cVeg) %>% 
  left_join(
    varnams %>% 
      select(modl, varn_cLeaf = cLeaf, varn_cVeg = cVeg),
    by = "modl"
  ) %>% 
  mutate_at(vars(starts_with("filn_")), ~str_replace(., ".nc", "")) %>% 
  mutate(dir = paste0("/cluster/home/bestocke/data/trendy/v8/")) %>% 
  mutate(cmd = paste("./calc_mass_fraction.sh", filn_cLeaf, filn_cVeg, dir))
```

Execute function using CDO.
```{r}
out <- purrr::map(as.list(df %>% pull(cmd)), ~system(., intern = TRUE))
```

### Root mass fraction

Determine for which models we have all the files (cWood and cVeg) to calculate wood mass fractions and create system command.
```{r}
modls_mr <- availvars %>% 
  filter(cRoot==1 & cVeg==1) %>% 
  pull(modl)

df <- filnams %>% 
  filter(modl %in% modls_mr) %>% 
  select(modl, filn_cRoot = cRoot, filn_cVeg = cVeg) %>% 
  left_join(
    varnams %>% 
      select(modl, varn_cRoot = cRoot, varn_cVeg = cVeg),
    by = "modl"
  ) %>% 
  mutate_at(vars(starts_with("filn_")), ~str_replace(., ".nc", "")) %>% 
  mutate(dir = paste0("/cluster/home/bestocke/data/trendy/v8/")) %>% 
  mutate(cmd = paste("./calc_mass_fraction.sh", filn_cRoot, filn_cVeg, dir))
```

Execute function using CDO.
```{r}
out <- purrr::map(as.list(df %>% pull(cmd)), ~system(., intern = TRUE))
```

### Root:shoot ratio

The root:shoot ratio can be calculated in different ways.
$$
r_{RS} = \frac{M_R}{M_L + M_W} = \frac{C_R}{C-C_R} = \frac{C-C_L-C_W}{C_L + C_W}
$$
The first one, based on mass fractions, requires the same files to be available as the second one. Thus, it's actually just two ways. Let's determine which models provide enough information to calculate it based on the second or third option.

Second option:
```{r}
modls_II <- availvars %>% 
  filter(cRoot==1 & cVeg==1) %>% 
  pull(modl)
```

Third option:
```{r}
modls_III <- availvars %>% 
  filter(cLeaf==1 & cWood==1 & cVeg==1) %>% 
  pull(modl)
```

There is no model available for III but not for II:
```{r}
!(modls_III %in% modls_II)
!(modls_II %in% modls_III)
```
Hence, let's just do it with `modls_II`:
```{r}
df <- filnams %>% 
  filter(modl %in% modls_II) %>% 
  select(modl, filn_cRoot = cRoot, filn_cVeg = cVeg) %>% 
  left_join(
    varnams %>% 
      select(modl, varn_cRoot = cRoot, varn_cVeg = cVeg),
    by = "modl"
  ) %>% 
  mutate_at(vars(starts_with("filn_")), ~str_replace(., ".nc", "")) %>% 
  mutate(dir = paste0("/cluster/home/bestocke/data/trendy/v8/")) %>% 
  mutate(cmd = paste("./calc_rootshoot_ratio.sh", filn_cRoot, filn_cVeg, dir))
```

Execute function using CDO.
```{r}
out <- purrr::map(as.list(df %>% pull(cmd)), ~system(., intern = TRUE))
```
