---
title: "SOC-litter input-biomass relationship in DGVMs"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
---

```{r include=FALSE}
library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(readr)
library(stringr)
library(rbeni)
```

## Approach

Does (steady-state) SOC storage, simulated in DGVMs, scale with litter inputs? And do litter inputs scale with biomass? To investigate this, we can look at the following relationships:

- The relative enhancement in steady-state SOC vs. the relative enhancement in NPP. $C^\ast$ is the steady-state SOC. It is given by the initial SOC stocks in simulations (spun up to steady state). 
$$
\frac{\Delta C^\ast}{C^\ast} / \frac{\Delta NPP}{NPP_\text{init}}
$$
- The relative enhancement in steady-state SOC vs. the relative enhancement in biomass ($B$).
$$
\frac{\Delta C^\ast}{C^\ast} / \frac{\Delta B}{B_\text{init}}
$$
$\Delta$ is the difference between end and beginning of a transient simulation covering the historical period, where only CO2 changes (hence, soil turnover rates should not change in a typical 1st order kinetics-based SOC model).

Our hypothesis is that these relationships follow the 1:1 line.

$\Delta C^\ast$ cannot be extracted directly from simulations. But it can be estimated from the imbalance between soil respiration and litter input before a steady state is reached:
$$
\tau \approx \frac{C(t)}{\text{NPP}(t) - \Delta C(t)}
$$

Combine this with $\Delta C^\ast=\tau \; \text{NPP}$, to get:
$$
C^\ast = \frac{C(t)}{1 - \frac{\Delta C(t)}{\text{NPP}(t)}}
$$
These variables are extracted from TRENDY S1 simulations as follows:

- $C(t)$: soil C at the end of the simulation (mean over 2008-2017)
- $\Delta C(t)$: Annual change in soil C during the last decade (2008-2017) (`filn_cSoil_change` - MUST BE DIVIDED BY 10!)
- NPP$(t)$: Annual mean NPP during the last decade (2008-2017)

- $\Delta C^\ast$: $C^\ast$ (as defined above) minus soil C during the first decade of the simulation (spun up to steady state)

The same is done with biomass ($B$).


## Process TRENDY v7 files

(Adjust path for `filnams` to select v8 or v7)

From downloaded TRENDY outputs (here, simulation S1: only CO2 is changing), get global fields of: 

1. Multi-annual mean NPP at the beginning of the simulation
2. Multi-annual mean NPP at the end of the simulation
3. Multi-annual mean SOC at the beginning of the simulation
4. Multi-annual mean SOC at the end of the simulation
5. Multi-annual mean biomass at the beginning of the simulation
6. Multi-annual mean biomass at the end of the simulation
7. Change in SOC over a multi-annual period at the end of the simulation (for $\Delta C(t)$)

Simulations should cover 1700-2017 (318 years = time step in NetCDF outputs). But not all do.

- Starting in year 1700: CABLE-POP, CLM5.0, JSBACH, LPX-Bern, LPJ-GUESS, OCN, ORCHIDEE-CNP, SDGVM
- Starting in year 1701: CLASS-CTEM, ISAM, JULES, SURFEX
- Starting in year 1900: DLEM
- Starting in year 1901: ORCHIDEE
- Starting in year 1860: VISIT

Peculiarities of outputs by model prohibiting their processing:

- LPJ: No S1 outputs
- DLEM: failed NPP processing
- LPX-Bern: failed NPP processing
- VISIT: No NPP output available

No separate output for pools (wood, leaf, root)

- CLASS-CTEM
- DLEM
- JSBACH
- JULES
- OCN
- SDGVM (only wood missing, could be derived by `cdo -O sub SDGVM_S1_cVeg.nc SDGVM_S1_cRoot.nc SDGVM_S1_cLeaf.nc`. cWood is then assumed zero: `cdo mulc,0 SDGVM_S1_cLeaf.nc SDGVM_S1_cWood.nc`, and `cdo -O chname,cLeaf,cWood SDGVM_S1_cWood.nc SDGVM_S1_cWood2.nc`, but is not!)
- SURFEX
- VISIT

### Process files: multi-annual mean

Determine for which models we have all the files (cWood and cVeg) to calculate wood mass fractions and create system command.
```{r}
# availvars <- read_csv( "./availvars_trendy_v8_S1.csv" )
# filnams <- read_csv( "filnams_trendy_v8_S1.csv" ) %>% 
filnams <- read_csv( "filnams_trendy_v7_S1.csv" ) %>% 
  setNames(c("modl", paste0("filn_", names(.)[-1])))

modls <- filnams %>% 
  filter(!is.na(filn_cSoil) & !is.na(filn_cVeg) & !is.na(filn_npp)) %>%

  ## for separate aboveground biomass
  filter(!is.na(filn_cWood) & !is.na(filn_cLeaf)) %>%
  
  ## remove models from analysis. See above for reasons for exclusion
  filter(!(modl %in% c("LPJ", "DLEM", "LPX-Bern", "VISIT"))) %>%
  
  filter(modl != "SDGVM") %>% 
  
  pull(modl)

df <- filnams %>% 
  
  ## filter based on above
  filter(modl %in% modls) %>%
  
  # mutate(dir = paste0("/cluster/home/bestocke/data/trendy/v8/")) %>% 
  # mutate(dir = paste0("~/data/trendy/v8/")) %>% 
  mutate(dir = paste0("~/data/trendy/v7/")) %>% 
  mutate_at(vars(starts_with("filn")), ~str_replace(., ".nc", "")) %>%
  
  ## get starting year
  left_join(read_csv("./startyear_trendy_v7_S1.csv"), by = "modl") %>% 
  rename(startyear_init = startyear, startyear_npp_init = startyear_npp) %>% 
  mutate(endyear_init = startyear_init + 9, endyear_npp_init = startyear_npp_init + 9) %>% 
    
  ## create CDO command to convert flux (per seconds) to annual total (per year)
  mutate(cmd_npp_units = paste("./get_tot_ann.sh", filn_npp, dir)) %>% 
  
  ## create CDO command to take mean across years at simulation start and end
  mutate(cmd_cVeg  = paste("./proc_trendy_soc_biomass.sh", filn_cVeg,  dir, startyear_init, endyear_init),
         cmd_cSoil = paste("./proc_trendy_soc_biomass.sh", filn_cSoil, dir, startyear_init, endyear_init),
         cmd_npp   = paste("./proc_trendy_soc_biomass.sh", filn_npp_ANN ,  dir, startyear_npp_init, endyear_npp_init),
         cmd_cLeaf = paste("./proc_trendy_soc_biomass.sh", filn_cLeaf, dir, startyear_init, endyear_init),
         cmd_cWood = paste("./proc_trendy_soc_biomass.sh", filn_cWood, dir, startyear_init, endyear_init),
         cmd_cRoot = paste("./proc_trendy_soc_biomass.sh", filn_cRoot, dir, startyear_init, endyear_init),
         cmd_dcSoil = paste("./proc_trendy_dsoc.sh", filn_cSoil, dir),
         cmd_dcVeg  = paste("./proc_trendy_dsoc.sh", filn_cVeg,  dir)
         ) %>% 
  
  rowwise() %>% 
  mutate(filn_cVeg_init  = paste0(filn_cVeg, "_INIT_MEAN"),
         filn_cVeg_final = paste0(filn_cVeg, "_FINAL_MEAN"),

         filn_cLeaf_init  = paste0(filn_cLeaf, "_INIT_MEAN"),
         filn_cLeaf_final = paste0(filn_cLeaf, "_FINAL_MEAN"),

         filn_cRoot_init  = paste0(filn_cRoot, "_INIT_MEAN"),
         filn_cRoot_final = paste0(filn_cRoot, "_FINAL_MEAN"),

         filn_cWood_init  = paste0(filn_cWood, "_INIT_MEAN"),
         filn_cWood_final = paste0(filn_cWood, "_FINAL_MEAN"),
         
         filn_cSoil_init  = paste0(filn_cSoil, "_INIT_MEAN"),
         filn_cSoil_final = paste0(filn_cSoil, "_FINAL_MEAN"),
         
         filn_cSoil_change = paste0(filn_cSoil, "_CHANGE"),
         
         filn_npp_init  = paste0(filn_npp_ANN, "_INIT_MEAN"),
         filn_npp_final = paste0(filn_npp_ANN, "_FINAL_MEAN")
         )
```

Execute function using CDO.
```{r eval=FALSE}
## convert NPP units (worked except for LPX and DLEM)
out <- purrr::map(as.list(df %>% pull(cmd_npp_units)), ~system(., intern = TRUE))

## get cVeg as a mean over the period at the beginning and the end of the simulations
out <- purrr::map(as.list(df %>% pull(cmd_cVeg)),   ~system(., intern = TRUE))

## get cSoil as a mean over the period at the beginning and the end of the simulations
out <- purrr::map(as.list(df %>% pull(cmd_cSoil)),  ~system(., intern = TRUE))

## get NPP as a mean over the period at the beginning and the end of the simulations
out <- purrr::map(as.list(df %>% pull(cmd_npp)),    ~system(., intern = TRUE))

## get additional fields needed to estimate steady-state for soil C
out <- purrr::map(as.list(df %>% pull(cmd_dcSoil)), ~system(., intern = TRUE))

## get additional fields needed to estimate steady-state for veg C
out <- purrr::map(as.list(df %>% pull(cmd_dcVeg)), ~system(., intern = TRUE))

## for remaining...
out <- purrr::map(as.list(df %>% pull(cmd_cLeaf)),  ~system(., intern = TRUE))
out <- purrr::map(as.list(df %>% pull(cmd_cWood)),  ~system(., intern = TRUE))
out <- purrr::map(as.list(df %>% pull(cmd_cRoot)),  ~system(., intern = TRUE))
```

## Collect data

... into tidy data frame

```{r}
source("R/collect_gdf_bymodl.R")

gdf <- purrr::map(

  as.list(seq(nrow(df))),

  ~collect_gdf_bymodl(

    df$modl[.],
    df$dir[.],

    df$filn_cSoil_init[.],
    df$filn_cSoil_final[.],

    df$filn_cVeg_init[.],
    df$filn_cVeg_final[.],

    df$filn_cLeaf_init[.],
    df$filn_cLeaf_final[.],

    df$filn_cRoot_init[.],
    df$filn_cRoot_final[.],

    df$filn_cWood_init[.],
    df$filn_cWood_final[.],

    df$filn_npp_init[.],
    df$filn_npp_final[.],

    df$filn_cSoil_change[.]
    )) %>% 
  
  bind_rows() %>%

  ## get aboveground biomass as sum of cWood and cLeaf
  mutate(cveg_ag_init = cleaf_init + cwood_init,
         cveg_ag_final = cleaf_final + cwood_final
         ) %>%

  group_by(modl) %>%
  nest()
```

Some tests.
```{r}
make_hist <- function(df){
  gg <- df %>% 
    mutate(dnpp = npp_final / npp_init) %>% 
    filter(dnpp > 0.2 & dnpp < 3) %>% 
    ggplot(aes(x = dnpp, y = ..density..)) +
    geom_histogram()
  return(gg)
}

gdf <- gdf %>% 
  mutate(gg_hist_dnpp = purrr::map(data, ~make_hist(.)))

testdf <- gdf$data[[1]] %>%
  mutate(dnpp = npp_final / npp_init)
```

<!-- Weird:  -->
<!-- - CABLE has constant NPP. -->

## Calculate relationships

```{r}
get_deltas <- function(df){
  df %>%
    mutate(dcveg = (cveg_final - cveg_init)/cveg_init,
           dcveg_ag = (cveg_ag_final - cveg_ag_init)/cveg_ag_init,
           dnpp = (npp_final - npp_init)/npp_init,
           dcsoil = (csoil_final - csoil_init)/csoil_init,
           dcsoil_star = (csoil_star_final - csoil_star_init)/csoil_star_init
           )
}

get_csoil_star <- function(df){
  df %>%
    mutate(csoil_star_init  = csoil_init) %>%
    # mutate(csoil_change = csoil_final - csoil_init) %>%   # overwriting what's read from file - XXX This was wrong XXX
    mutate(csoil_change = csoil_change / 10.0) %>%
    mutate(csoil_star_final = csoil_final / (1.0 - (csoil_change/10.0)/npp_final ))
}

gdf <- gdf %>%
  mutate(data = purrr::map(data, ~get_csoil_star(.))) %>%
  mutate(data = purrr::map(data, ~get_deltas(.)))
```

Plot relationships:

### Biomass vs. NPP change

```{r}
## just one model
modobs <- gdf$data[[2]] %>%
  dplyr::filter(dnpp < 2 & dnpp > -2 & dcveg < 2 & dcveg > -2) %>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg)) %>%
  analyse_modobs2("dnpp", "dcveg", type = "heat", plot_subtitle = FALSE, plot_linmod = FALSE)
modobs$gg +
  xlim(c(-0.5,1)) + ylim(c(-0.5,1)) +
  labs(title = gdf$modl[[2]])

## all models pooled
gdf %>%
  unnest(data) %>%
  dplyr::filter(dnpp < 2 & dnpp > -2 & dcveg < 2 & dcveg > -2) %>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg)) %>%
  ggplot(aes(x = dnpp, y = dcveg)) +
  stat_density_2d(aes(fill = after_stat(level)), geom = "polygon") +
  theme_classic() +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  scale_fill_gradientn(colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5))

## all models separately (facet_grid)
gdf %>%
  dplyr::filter(modl != "CABLE-POP") %>%
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      dplyr::filter(modl != "CABLE-POP") %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = 4420))) %>% # smallest set here
      unnest(data) %>%
      mutate(modl = "ALL")
  ) %>%
  dplyr::filter(dnpp>-0.9) %>% # remove points where veg collapses
  dplyr::filter(dnpp < 2 & dnpp > -2 & dcveg < 2 & dcveg > -2) %>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg)) %>%
  ggplot(aes(x = dnpp, y = dcveg)) +
  stat_density_2d(aes(fill = after_stat(nlevel)), geom = "polygon") +
  theme_classic() +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  scale_fill_gradientn(colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5),
                       guide = "legend") +
  facet_wrap(. ~ modl, nrow = 3,  scales = "free") +
  xlim(-0.1, 0.9) + ylim(-0.1, 0.9) +
  theme(legend.position = "none",
        strip.background = element_blank())

ggsave("fig/cveg_dnpp_trendy_s1_bymodl.pdf", width = 9, height = 7)

# modobs <- gdf %>%
#   unnest(data) %>%
#   dplyr::filter(dnpp < 2 & dnpp > -2 & dcveg < 2 & dcveg > -2) %>%
#   dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg)) %>%
#   analyse_modobs2("dnpp", "dcveg", type = "hex", plot_subtitle = FALSE, plot_linmod = FALSE)
# modobs$gg +
#   xlim(c(-0.5,1)) + ylim(c(-0.5,1)) +
#   labs(title = "All models pooled")
```

### SOC vs. NPP change

```{r}
## one model
modobs <- gdf$data[[2]] %>%
  dplyr::filter(dnpp < 2 & dnpp > -2 & dcsoil_star < 2 & dcsoil_star > -2) %>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dcsoil_star), !is.infinite(dnpp), !is.infinite(dcsoil_star)) %>%
  analyse_modobs2("dnpp", "dcsoil_star", type = "heat", plot_subtitle = FALSE, plot_linmod = FALSE)
modobs$gg +
  xlim(c(-0.5,1)) + ylim(c(-0.5,1)) +
  labs(title = gdf$modl[[2]])

## all models pooled
modobs <- gdf %>%
  unnest(data) %>%
  dplyr::filter(dnpp < 2 & dnpp > -2 & dcsoil_star < 2 & dcsoil_star > -2) %>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dcsoil_star), !is.infinite(dnpp), !is.infinite(dcsoil_star)) %>%
  analyse_modobs2("dnpp", "dcsoil_star", type = "hex", plot_subtitle = FALSE, plot_linmod = FALSE)
modobs$gg +
  xlim(c(-0.1,0.7)) + ylim(c(-0.1,0.7)) +
  labs(title = "All models pooled")

## cSoil* vs. NPP, all models separately (facet_grid)
gdf %>%
  dplyr::filter(modl != "CABLE-POP") %>%
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      dplyr::filter(modl != "CABLE-POP") %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = 4420))) %>% # smallest set here
      unnest(data) %>%
      mutate(modl = "ALL")
  ) %>%
  dplyr::filter(dcsoil_star < 2 & dcsoil_star > -2 & dnpp < 2 & dnpp > -2) %>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dnpp), !is.infinite(dnpp), !is.infinite(dnpp)) %>%
  ggplot(aes(x = dnpp, y = dcsoil_star)) +
  stat_density_2d(aes(fill = after_stat(nlevel)), geom = "polygon") +
  theme_classic() +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  scale_fill_gradientn(colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5),
                       guide = "legend") +
  facet_wrap(. ~ modl, nrow = 3,  scales = "free") +
  xlim(-0.3, 1.2) + ylim(-0.3, 1.2) +
  labs(x = expression(paste(Delta, "NPP", "/NPP")),
       y = expression(paste(Delta, "C"[soil], "/C"[soil]))) +
  theme(legend.position = "none",
        strip.background = element_blank())

ggsave("fig/csoil_dnpp_trendy_s1_bymodl.pdf", width = 9, height = 7)
```

### SOC vs. biomass change

```{r}
# # one model
# modobs <- gdf$data[[2]] %>%
#   dplyr::filter(dcveg < 2 & dcveg > -2 & dcsoil_star < 2 & dcsoil_star > -2) %>%
#   dplyr::filter(!is.nan(dcveg), !is.nan(dcsoil_star), !is.infinite(dcveg), !is.infinite(dcsoil_star)) %>%
#   analyse_modobs2("dcveg", "dcsoil_star", type = "heat", plot_subtitle = FALSE, plot_linmod = FALSE)
# modobs$gg +
#   xlim(c(-0.5,1)) + ylim(c(-0.5,1)) +
#   labs(title = gdf$modl[[2]])

## cSoil*, all models pooled
gdf %>%
  unnest(data) %>%
  dplyr::filter(dcsoil_star < 2 & dcsoil_star > -2 & dcveg < 2 & dcveg > -2) %>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg)) %>%
  
  ggplot(aes(x = dcveg, y = dcsoil_star)) +
  
  stat_density_2d(aes(fill = after_stat(nlevel)), geom = "polygon") +
  theme_classic() +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  scale_fill_gradientn(colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5),
                       guide = "legend") +
  xlim(-0.1, 1) + ylim(-0.1, 1) +
  labs(x = expression(paste(Delta, "C"[veg], "/C"[veg])),
       y = expression(paste(Delta, "C"[soil], "/C"[soil])))
  # theme(legend.position = "none",
  #       strip.background = element_blank())

## cSoil*, all models separately (facet_grid)
gdf %>%
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = 1956))) %>% # smallest set here
      unnest(data) %>%
      mutate(modl = "ALL")
  ) %>%
  dplyr::filter(dcsoil_star < 2 & dcsoil_star > -2 & dcveg < 2 & dcveg > -2) %>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg)) %>%
  
  ggplot(aes(x = dcveg, y = dcsoil_star)) +
  stat_density_2d(aes(fill = after_stat(nlevel)), geom = "polygon") +
  theme_classic() +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  scale_fill_gradientn(colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5),
                       guide = "legend") +
  facet_wrap(. ~ modl, nrow = 3,  scales = "free") +
  xlim(-0.1, 1) + ylim(-0.1, 1) +
  labs(x = expression(paste(Delta, "C"[veg], "/C"[veg])),
       y = expression(paste(Delta, "C"[soil], "/C"[soil]))) +
  theme(legend.position = "none",
        strip.background = element_blank())

ggsave("fig/csoil_cveg_trendy_s1_bymodl.pdf", width = 9, height = 7)

## cSoil, all models separately (facet_grid)
gdf %>%
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = 4420))) %>% # smallest set here
      unnest(data) %>%
      mutate(modl = "ALL")
  ) %>%
  dplyr::filter(dcsoil < 2 & dcsoil > -2 & dcveg < 2 & dcveg > -2) %>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg)) %>%
  
  ggplot(aes(x = dcveg, y = dcsoil)) +
  stat_density_2d(aes(fill = after_stat(nlevel)), geom = "polygon") +
  theme_classic() +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  scale_fill_gradientn(colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5),
                       guide = "legend") +
  facet_wrap(. ~ modl, nrow = 3,  scales = "free") +
  xlim(-0.1, 0.9) + ylim(-0.1, 0.9) +
  theme(legend.position = "none",
        strip.background = element_blank())
```

### SOC vs. aboveground biomass change

```{r}
# # one model
# modobs <- gdf$data[[3]] %>%
#   dplyr::filter(dcveg_ag < 2 & dcveg_ag > -2 & dcsoil_star < 2 & dcsoil_star > -2) %>%
#   dplyr::filter(!is.nan(dcveg_ag), !is.nan(dcsoil_star), !is.infinite(dcveg_ag), !is.infinite(dcsoil_star)) %>%
#   analyse_modobs2("dcveg_ag", "dcsoil_star", type = "heat", plot_subtitle = FALSE, plot_linmod = FALSE)
# modobs$gg +
#   xlim(c(-0.5,1)) + ylim(c(-0.5,1)) +
#   labs(title = gdf$modl[[3]])

## all models pooled
gdf %>%
  unnest(data) %>%
  dplyr::filter(dcveg_ag < 2 & dcveg_ag > -2 & dcsoil_star < 2 & dcsoil_star > -2) %>%
  dplyr::filter(!is.nan(dcveg_ag), !is.nan(dcsoil_star), !is.infinite(dcveg_ag), !is.infinite(dcsoil_star)) %>%

  ggplot(aes(x = dcveg_ag, y = dcsoil_star)) +
  stat_density_2d(aes(fill = after_stat(nlevel)), geom = "polygon") +
  theme_classic() +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  geom_vline(xintercept = 0, color = "grey50", size = 0.2) +
  geom_hline(yintercept = 0, color = "grey50", size = 0.2) +
  scale_fill_gradientn(colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5),
                       guide = "legend") +
  xlim(-0.5, 1.5) + ylim(-0.5, 1.5) +
  labs(x = expression(paste(Delta, "C"[ag-veg], "/C"[ag-veg])),
       y = expression(paste(Delta, "C"[soil], "/C"[soil]))) +
  theme(legend.position = "none",
        strip.background = element_blank())


## cSoil*, all models separately (facet_grid)
gdf %>%
  unnest(data) %>%
  # bind_rows(
  #   .,
  #   gdf %>%
  #     mutate(n = purrr::map_int(data, ~nrow(.))) %>%
  #     mutate(data = purrr::map(data, ~slice_sample(., n = 4420))) %>% # smallest set here
  #     unnest(data) %>%
  #     mutate(modl = "ALL")
  # ) %>%
  dplyr::filter(dcsoil_star < 2 & dcsoil_star > -2 & dcveg_ag < 2 & dcveg_ag > -2) %>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg_ag), !is.infinite(dnpp), !is.infinite(dcveg_ag)) %>%
  ggplot(aes(x = dcveg_ag, y = dcsoil_star)) +
  stat_density_2d(aes(fill = after_stat(nlevel)), geom = "polygon") +
  theme_classic() +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  geom_vline(xintercept = 0, color = "grey50", size = 0.2) +
  geom_hline(yintercept = 0, color = "grey50", size = 0.2) +
  scale_fill_gradientn(colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5),
                       guide = "legend") +
  facet_wrap(. ~ modl, nrow = 3,  scales = "free") +
  xlim(-0.2, 1.0) + ylim(-0.2, 1.0) +
  labs(x = expression(paste(Delta, "C"[ag-veg], "/C"[ag-veg])),
       y = expression(paste(Delta, "C"[soil], "/C"[soil]))) +
  theme(legend.position = "none",
        strip.background = element_blank())

ggsave("fig/csoil_cveg_ag_trendy_s1_bymodl.pdf", width = 7, height = 7)


## cSoil*, all models pooled
gdf %>%

  ## sample 1957 points from each model (corresponds to the number of gridcell of the model with coarsest res.)
  mutate(data = purrr::map(data, ~slice_sample(., n = 1957))) %>% # smallest set here
  unnest(data) %>%
  mutate(modl = "ALL") %>%

  dplyr::filter(dcsoil_star < 2 & dcsoil_star > -2 & dcveg_ag < 2 & dcveg_ag > -2) %>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg_ag), !is.infinite(dnpp), !is.infinite(dcveg_ag)) %>%
  ggplot(aes(x = dcveg_ag, y = dcsoil_star)) +
  stat_density_2d(aes(fill = after_stat(nlevel)), geom = "polygon") +
  theme_classic() +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  geom_vline(xintercept = 0, color = "grey50", size = 0.2) +
  geom_hline(yintercept = 0, color = "grey50", size = 0.2) +
  scale_fill_gradientn(colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5),
                       guide = "legend") +
  # facet_wrap(. ~ modl, nrow = 3,  scales = "free") +
  xlim(-0.5, 1.5) + ylim(-0.5, 1.5) +
  labs(x = expression(paste(Delta, "C"[ag-veg], "/C"[ag-veg])),
       y = expression(paste(Delta, "C"[soil], "/C"[soil]))) +
  theme(legend.position = "none",
        strip.background = element_blank())

ggsave("fig/csoil_cveg_ag_trendy_s1.pdf", width = 7, height = 7)

# ## Cesar's figure
# df_cesar <- read_csv("./data/model_data_mean_long.csv")
#
# obs.mods <- ggplot(df_cesar, aes(make_pct(Cabove), make_pct(Csoil), colour=type)) +
#   geom_smooth(size=.5,data=filter(df_cesar,type=="modeled"), method = "lm", se=FALSE, color="#e41a1c", formula = y ~ x) +
#   stat_smooth(size=.5,data=filter(df_cesar,type=="observed"), method = "lm", se=FALSE, color="#377eb8", formula = y ~ poly(x, 3)) + # orthogonal polynomials, in this case with a 3rd order
#   geom_errorbar(aes(ymin=make_pct(Csoil)-make_pct(SE.soil), ymax=make_pct(Csoil)+make_pct(SE.soil)), width=1,size=.3,colour="black") +
#   geom_point(aes(shape=Site),size=1.5,fill="white",alpha=.7) +
#   scale_shape_manual(values=c(21:25,4)) + guides(colour=guide_legend(title=NULL)) +
#   scale_colour_manual(values=c("#e41a1c","#377eb8")) +
#   guides(col = guide_legend(title = NULL,order = 0),
#          shape = guide_legend(title = "Experiment",order = 1)) +
#   labs(x=expression(paste(CO[2]," effect on biomass carbon (%)", sep="")),
#        y=expression(paste(CO[2]," effect on soil carbon (%)", sep=""))) +
#   theme_cowplot(font_size=8) +
#   theme(legend.position = c(0.7, 0.7),
#         legend.spacing.y = unit(2, "pt"))


## cSoil, all models separately (facet_grid)
gdf %>%
  dplyr::filter(modl != "CABLE-POP") %>%
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      dplyr::filter(modl != "CABLE-POP") %>%
      mutate(n = purrr::map_int(data, ~nrow(.))) %>%
      mutate(data = purrr::map(data, ~slice_sample(., n = 4420))) %>% # smallest set here
      unnest(data) %>%
      mutate(modl = "ALL")
  ) %>%
  dplyr::filter(dcsoil < 2 & dcsoil > -2 & dcveg_ag < 2 & dcveg_ag > -2) %>%
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg_ag), !is.infinite(dnpp), !is.infinite(dcveg_ag)) %>%
  ggplot(aes(x = dcveg_ag, y = dcsoil)) +
  stat_density_2d(aes(fill = after_stat(nlevel)), geom = "polygon") +
  theme_classic() +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  scale_fill_gradientn(colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5),
                       guide = "legend") +
  facet_wrap(. ~ modl, nrow = 3,  scales = "free") +
  xlim(-0.1, 0.9) + ylim(-0.1, 0.9) +
  theme(legend.position = "none",
        strip.background = element_blank())
```