---
title: "SOC-litter input-biomass relationship in DGVMs"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
---

```{r include=FALSE}
library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(readr)
library(stringr)
library(rbeni)
```

## Approach

Does (steady-state) SOC storage, simulated in DGVMs, scale with litter inputs? And do litter inputs scale with biomass? To investigate this, can look at the following relationships:

- $\frac{\Delta C^\ast}{C^\ast}$ vs. $\frac{\Delta NPP}{NPP_\text{init}}$: The relative enhancement in steady-state SOC vs. the relative enhancement in NPP. $C^\ast$ is the steady-state SOC. It is given by the initial SOC stocks in simulations (spun up to steady state). 
- $\frac{\Delta C^\ast}{C^\ast}$ vs. $\frac{\Delta B}{B_\text{init}}$: The relative enhancement in steady-state SOC vs. the relative enhancement in biomass ($B$).

$\Delta$ is the difference between end and beginning of a transient simulation covering the historical period, where only CO2 changes (hence, soil turnover rates should not change in a typical 1st order kinetics-based SOC model).

Our hypothesis is that these relationships follow the 1:1 line.

$\Delta C^\ast$ cannot be extracted directly from simulations. But it can be estimated from the imbalance between soil respiration and litter input before a steady state is reached:

$$
\tau \approx \frac{C(t)}{\text{NPP}(t) - \Delta C(t)}
$$

Combine this with $\Delta C^\ast=\tau \; \text{NPP}$, to get:
$$
C^\ast = \frac{C(t)}{1 - \frac{\Delta C(t)}{\text{NPP}(t)}}
$$

The same is done with biomass ($B$). For NPP, we can assume that steady-state is immediately reached (disregarding expansion of leaf area).

## Process TRENDY v8 files

From downloaded TRENDY outputs (here, simulation S2: only CO2 is changing), get global fields of: 

1. Multi-annual mean NPP at the beginning of the simulation
2. Multi-annual mean NPP at the end of the simulation
3. Multi-annual mean SOC at the beginning of the simulation
4. Multi-annual mean SOC at the end of the simulation
5. Multi-annual mean biomass at the beginning of the simulation
6. Multi-annual mean biomass at the end of the simulation
7. Change in SOC over a multi-annual period at the end of the simulation (for $\Delta C(t)$)

### Process files: multi-annual mean

Determine for which models we have all the files (cWood and cVeg) to calculate wood mass fractions and create system command.
```{r}
# availvars <- read_csv( "./availvars_trendy_v8_S1.csv" )
filnams <- read_csv( "filnams_trendy_v8_S1.csv" ) %>% 
  setNames(c("modl", paste0("filn_", names(.)[-1])))

modls <- filnams %>% 
  filter(!is.na(filn_cSoil) & !is.na(filn_cVeg) & !is.na(filn_npp)) %>%
  pull(modl)

df <- filnams %>% 
  filter(modl %in% modls) %>%
  # mutate(dir = paste0("/cluster/home/bestocke/data/trendy/v8/")) %>% 
  mutate(dir = paste0("~/data/trendy/v8/")) %>% 
  mutate_at(1:10, ~str_replace(., ".nc", "")) %>% 
  mutate(cmd_cVeg  = paste("./proc_trendy_soc_biomass.sh", filn_cVeg,  dir),
         cmd_npp   = paste("./proc_trendy_soc_biomass.sh", filn_npp ,  dir),
         cmd_cSoil = paste("./proc_trendy_soc_biomass.sh", filn_cSoil, dir),
         cmd_dcSoil = paste("./proc_trendy_dsoc.sh", filn_cSoil, dir)
         ) %>% 
  rowwise() %>% 
  mutate(filn_cVeg_init  = paste0(filn_cVeg, "_INIT_MEAN"),
         filn_cVeg_final = paste0(filn_cVeg, "_FINAL_MEAN"),
         filn_cSoil_init  = paste0(filn_cSoil, "_INIT_MEAN"),
         filn_cSoil_final = paste0(filn_cSoil, "_FINAL_MEAN"),
         filn_cSoil_change = paste0(filn_cSoil, "_CHANGE"),
         filn_npp_init  = paste0(filn_npp, "_INIT_MEAN"),
         filn_npp_final = paste0(filn_npp, "_FINAL_MEAN"))
```

Execute function using CDO.
```{r}
out <- purrr::map(as.list(df %>% pull(cmd_cVeg)),  ~system(., intern = TRUE))
out <- purrr::map(as.list(df %>% pull(cmd_npp)),   ~system(., intern = TRUE))
out <- purrr::map(as.list(df %>% pull(cmd_cSoil)), ~system(., intern = TRUE))
out <- purrr::map(as.list(df %>% pull(cmd_dcSoil)), ~system(., intern = TRUE))
```

## Collect data 

... into tidy data frame

```{r}
source("R/collect_gdf_bymodl.R")
gdf <- purrr::map(as.list(seq(nrow(df))),
  ~collect_gdf_bymodl(
    df$modl[.], 
    df$dir[.], 
    df$filn_cSoil_init[.], 
    df$filn_cSoil_final[.], 
    df$filn_cVeg_init[.], 
    df$filn_cVeg_final[.], 
    df$filn_npp_init[.], 
    df$filn_npp_final[.], 
    df$filn_cSoil_change[.]) 
    )
names(gdf) <- df$modl
gdf <- gdf %>% 
  bind_rows(.id = "modl") %>% 
  group_by(modl) %>% 
  nest()
```

Some tests.
```{r}
testdf <- gdf$data[[2]] %>% 
  mutate(dnpp = npp_final / npp_init)

testdf$dnpp %>% quantile(., probs = c(0.1, 0.5, 0.9), na.rm=T)
```

Weird: 
- CABLE has constant NPP.

## Calculate relationships

```{r}
get_deltas <- function(df){
  df %>% 
    mutate(dcveg = (cveg_final - cveg_init)/cveg_init,
           dnpp = (npp_final - npp_init)/npp_init,
           dcsoil = (csoil_final - csoil_init)/csoil_init,
           dcsoil_star = (csoil_star_final - csoil_star_init)/csoil_star_init
           )
}

get_csoil_star <- function(df){
  df %>% 
    mutate(csoil_star_init  = csoil_init) %>% 
    # mutate(csoil_change = csoil_final - csoil_init) %>%   # overwriting what's read from file
    mutate(csoil_star_final = csoil_final / (1.0 - csoil_change/npp_final ))
}

gdf <- gdf %>% 
  mutate(data = purrr::map(data, ~get_csoil_star(.))) %>% 
  mutate(data = purrr::map(data, ~get_deltas(.)))
```

Plot relationships:

<!-- ### Biomass vs. NPP change -->

<!-- ```{r} -->
<!-- ## just one model -->
<!-- modobs <- gdf$data[[2]] %>%  -->
<!--   dplyr::filter(dnpp < 2 & dnpp > -2 & dcveg < 2 & dcveg > -2) %>%  -->
<!--   dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg)) %>%  -->
<!--   analyse_modobs2("dnpp", "dcveg", type = "heat", plot_subtitle = FALSE, plot_linmod = FALSE) -->
<!-- modobs$gg + -->
<!--   xlim(c(-0.5,1)) + ylim(c(-0.5,1)) + -->
<!--   labs(title = gdf$modl[[2]]) -->

<!-- ## all models pooled -->
<!-- gdf %>% -->
<!--   unnest(data) %>% -->
<!--   dplyr::filter(dnpp < 2 & dnpp > -2 & dcveg < 2 & dcveg > -2) %>%  -->
<!--   dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg)) %>%  -->
<!--   ggplot(aes(x = dnpp, y = dcveg)) + -->
<!--   stat_density_2d(aes(fill = after_stat(level)), geom = "polygon") +  -->
<!--   theme_classic() + -->
<!--   geom_abline(intercept=0, slope=1, linetype="dotted") + -->
<!--   scale_fill_gradientn(colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5)) -->

<!-- ## all models separately (facet_grid) -->
<!-- gdf %>% -->
<!--   dplyr::filter(modl != "CABLE-POP") %>%  -->
<!--   unnest(data) %>% -->
<!--   bind_rows( -->
<!--     ., -->
<!--     gdf %>% -->
<!--       dplyr::filter(modl != "CABLE-POP") %>%  -->
<!--       unnest(data) %>% -->
<!--       mutate(modl = "ALL") -->
<!--   ) %>%  -->
<!--   dplyr::filter(dnpp>-0.9) %>% # remove points where veg collapses -->
<!--   dplyr::filter(dnpp < 2 & dnpp > -2 & dcveg < 2 & dcveg > -2) %>%  -->
<!--   dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg)) %>%  -->
<!--   ggplot(aes(x = dnpp, y = dcveg)) + -->
<!--   stat_density_2d(aes(fill = after_stat(nlevel)), geom = "polygon") +  -->
<!--   theme_classic() + -->
<!--   geom_abline(intercept=0, slope=1, linetype="dotted") + -->
<!--   scale_fill_gradientn(colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5), -->
<!--                        guide = "legend") + -->
<!--   facet_wrap(. ~ modl, nrow = 3,  scales = "free") + -->
<!--   xlim(-0.1, 0.9) + ylim(-0.1, 0.9) +  -->
<!--   theme(legend.position = "none") -->
<!--   # scale_fill_binned(colors = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5)) + -->
<!--   # guides(fill = guide_colorsteps()) -->
<!--   # scale_fill_binned(type = "viridis") -->

<!-- # modobs <- gdf %>%  -->
<!-- #   unnest(data) %>%  -->
<!-- #   dplyr::filter(dnpp < 2 & dnpp > -2 & dcveg < 2 & dcveg > -2) %>%  -->
<!-- #   dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg)) %>%  -->
<!-- #   analyse_modobs2("dnpp", "dcveg", type = "hex", plot_subtitle = FALSE, plot_linmod = FALSE) -->
<!-- # modobs$gg + -->
<!-- #   xlim(c(-0.5,1)) + ylim(c(-0.5,1)) + -->
<!-- #   labs(title = "All models pooled") -->
<!-- ``` -->

<!-- ### SOC vs. NPP change -->



<!-- ```{r} -->
<!-- ## one model -->
<!-- modobs <- gdf$data[[2]] %>%  -->
<!--   dplyr::filter(dnpp < 2 & dnpp > -2 & dcsoil_star < 2 & dcsoil_star > -2) %>%  -->
<!--   dplyr::filter(!is.nan(dnpp), !is.nan(dcsoil_star), !is.infinite(dnpp), !is.infinite(dcsoil_star)) %>%  -->
<!--   analyse_modobs2("dnpp", "dcsoil_star", type = "heat", plot_subtitle = FALSE, plot_linmod = FALSE) -->
<!-- modobs$gg + -->
<!--   xlim(c(-0.5,1)) + ylim(c(-0.5,1)) + -->
<!--   labs(title = gdf$modl[[2]]) -->
<!-- ``` -->

<!-- ### SOC vs. biomass change -->

```{r}
# one model
modobs <- gdf$data[[2]] %>% 
  dplyr::filter(dcveg < 2 & dcveg > -2 & dcsoil_star < 2 & dcsoil_star > -2) %>% 
  dplyr::filter(!is.nan(dcveg), !is.nan(dcsoil_star), !is.infinite(dcveg), !is.infinite(dcsoil_star)) %>% 
  analyse_modobs2("dcveg", "dcsoil_star", type = "heat", plot_subtitle = FALSE, plot_linmod = FALSE)
modobs$gg +
  xlim(c(-0.5,1)) + ylim(c(-0.5,1)) +
  labs(title = gdf$modl[[2]])

## all models pooled
modobs <- gdf %>% 
  unnest(data) %>% 
  dplyr::filter(dcveg < 2 & dcveg > -2 & dcsoil_star < 2 & dcsoil_star > -2) %>% 
  dplyr::filter(!is.nan(dcveg), !is.nan(dcsoil_star), !is.infinite(dcveg), !is.infinite(dcsoil_star)) %>% 
  analyse_modobs2("dcveg", "dcsoil_star", type = "hex", plot_subtitle = FALSE, plot_linmod = FALSE)
modobs$gg +
  xlim(c(-0.5,1)) + ylim(c(-0.5,1)) +
  labs(title = "All models pooled")

## all models separately (facet_grid)
gdf %>%
  dplyr::filter(modl != "CABLE-POP") %>% 
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      dplyr::filter(modl != "CABLE-POP") %>% 
      unnest(data) %>%
      mutate(modl = "ALL")
  ) %>% 
  dplyr::filter(dcsoil_star < 2 & dcsoil_star > -2 & dcveg < 2 & dcveg > -2) %>% 
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg)) %>% 
  ggplot(aes(x = dcveg, y = dcsoil_star)) +
  stat_density_2d(aes(fill = after_stat(nlevel)), geom = "polygon") + 
  theme_classic() +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  scale_fill_gradientn(colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5)) +
  facet_wrap(. ~ modl, nrow = 2,  scales = "free")

## cSoil*, all models separately (facet_grid)
gdf %>%
  dplyr::filter(modl != "CABLE-POP") %>% 
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      dplyr::filter(modl != "CABLE-POP") %>% 
      unnest(data) %>%
      mutate(modl = "ALL")
  ) %>% 
  dplyr::filter(dcsoil_star < 2 & dcsoil_star > -2 & dcveg < 2 & dcveg > -2) %>% 
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg)) %>% 
  ggplot(aes(x = dcveg, y = dcsoil_star)) +
  stat_density_2d(aes(fill = after_stat(nlevel)), geom = "polygon") + 
  theme_classic() +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  scale_fill_gradientn(colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5),
                       guide = "legend") +
  facet_wrap(. ~ modl, nrow = 3,  scales = "free") +
  xlim(-0.1, 0.9) + ylim(-0.1, 0.9) + 
  theme(legend.position = "none",
        strip.background = element_blank())

## cSoil, all models separately (facet_grid)
gdf %>%
  dplyr::filter(modl != "CABLE-POP") %>% 
  unnest(data) %>%
  bind_rows(
    .,
    gdf %>%
      dplyr::filter(modl != "CABLE-POP") %>% 
      unnest(data) %>%
      mutate(modl = "ALL")
  ) %>% 
  dplyr::filter(dcsoil < 2 & dcsoil > -2 & dcveg < 2 & dcveg > -2) %>% 
  dplyr::filter(!is.nan(dnpp), !is.nan(dcveg), !is.infinite(dnpp), !is.infinite(dcveg)) %>% 
  ggplot(aes(x = dcveg, y = dcsoil)) +
  stat_density_2d(aes(fill = after_stat(nlevel)), geom = "polygon") + 
  theme_classic() +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  scale_fill_gradientn(colours = colorRampPalette( c("gray65", "navy", "red", "yellow"))(5),
                       guide = "legend") +
  facet_wrap(. ~ modl, nrow = 3,  scales = "free") +
  xlim(-0.1, 0.9) + ylim(-0.1, 0.9) + 
  theme(legend.position = "none",
        strip.background = element_blank())
```